---
title: "Lab 5"
author: "Erin Cagle"
date: "2023-09-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
1.  Load the `data.table` (and the `dtplyr` and `dplyr` packages if you plan to work with those).

    ```{r}
    library(data.table)
    library(dtplyr)
    library(dplyr)
    library(ggplot2)
    ```
2.  Load the met data from https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz, and also the station data. For the later, you can use the code we used during lecture to pre-process the stations data:

    ```{r}
    if (!file.exists("met_all.gz"))
      download.file(
        url = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz",
        destfile = "met_all.gz",
        method   = "libcurl",
        timeout  = 60
        )
    met <- data.table::fread("met_all.gz")
    ``````{r stations-data, eval = FALSE}

    # Download the data
stations <- fread("ftp://ftp.ncdc.noaa.gov/pub/data/noaa/isd-history.csv")
stations[, USAF := as.integer(USAF)]

# Dealing with NAs and 999999
stations[, USAF   := fifelse(USAF == 999999, NA_integer_, USAF)]
stations[, CTRY   := fifelse(CTRY == "", NA_character_, CTRY)]
stations[, STATE  := fifelse(STATE == "", NA_character_, STATE)]

# Selecting the three relevant columns, and keeping unique records
stations <- unique(stations[, list(USAF, CTRY, STATE)])

# Dropping NAs
stations <- stations[!is.na(USAF)]

# Removing duplicates
stations[, n := 1:.N, by = .(USAF)]
stations <- stations[n == 1,][, n := NULL]

#3. Merge the data as we did during the lecture.
stat <- merge(
  # Data
  x     = met,      
  y     = stations, 
  # List of variables to match
  by.x  = "USAFID",
  by.y  = "USAF", 
  # Which obs to keep?
  all.x = TRUE,      
  all.y = FALSE
  )
head(stat[, list(USAFID, WBAN, STATE)], n = 4)
```
## Question 1: Representative station for the US

What is the median station in terms of temperature, wind speed, and atmospheric pressure? Look for the three weather stations that best represent continental US using the `quantile()` function. Do these three coincide?

```{r}
stat_med <- stat[, .(
    temp_quan      = quantile(temp, na.rm=TRUE),
    wind.sp_quan   = quantile(wind.sp, na.rm=TRUE),
    atm.press_quan = quantile(atm.press, na.rm = TRUE)
    )]

# in_names  <- c("wind.sp", "temp", "atm.press")
# out_names <- paste0(in_names, "_med")
# stat_med <- stat[,
#   setNames(lapply(.SD, quantile, na.rm = TRUE), out_names),
#   .SDcols = in_names
#   ]

```

Median temp is 23.5, median wind speed